---
title: "Team Project 2"
author: "Team 22"
date: "4/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r}
library(dplyr)
library(ggplot2)
library(tidyr)
```

``` {r}
# take a long time and require 8 Gigabyte of spare RAM.

loan <- read.csv("loan.csv")


```


``` {r}
colSums(is.na(loan))

```


``` {r}
# delete any column that has more than 100000 NAs

threshold = 0.3

loan <- loan[, colSums(is.na(loan)) < nrow(loan) * threshold]

```


``` {r}
colSums(is.na(loan))
```


``` {r}
loan <- na.omit(loan )
colSums(is.na(loan))
```

### add a target variable on loan status, that divide 10 loan status into 2 category, making it a binary variable

``` {r}
#unique(loan$loan_status)

table(loan$loan_status)
barplot(table(loan$loan_status))


loan$target <- as.factor(ifelse(loan$loan_status %in% c("Charged Off",
                    "Default",
                    "Does not meet the credit policy. Status:Charged Off",
                    "In Grace Period", 
                    "Default Receiver", 
                    "Late (16-30 days)",
                    "Late (31-120 days)"), 0, ifelse(loan$loan_status %in% c("Fully Paid"), 1, 2)))
table(loan$target)

```

### filter out loans that are still "current", which corresponds to target = 2

``` {r}
cleaned_loan <- loan %>% filter(target != 2)
table(cleaned_loan$target)

```

### test the baseline of predicting all 1 and using the provided grade column

``` {r}
cleaned_loan$target <- factor(cleaned_loan$target)
(baseline_grade <- table(cleaned_loan$grade, cleaned_loan$target))

```

``` {r}
(all_1_acc <- nrow(cleaned_loan[which(as.integer(cleaned_loan$target) != 1),])/nrow(cleaned_loan))
(all_1_type2err <- 1)

(grade_acc <- sum(baseline_grade[c("A", "B", "C"),2])/sum(baseline_grade))
(grade_type2err <- sum(baseline_grade[c("D", "E", "F", "G"), 1])/sum(baseline_grade[, 1]))



```
  
Predicting all 1 results in an accuracy of 0.7734 and a Type 2 error of 1. We cannot predict any of the default.  
Using the provided grade column as indicator we manage to get a accuracy of 0.61 while predicting 43% out of all defaults. Noted that we can change the threshold for grade to adjust the trade-off.  
    
Our purpose is to develop a model that can minimize type 2 error, which in other words mean we want to be able to predict as many bad loans as possible, while maintaing a resonable level of accuracy.

Now we proceed to filter out variables that are unsuable such as the employee title because everyone has his or her own unique employee title, and hardship-related variables since there is no way we would know those when the loan originates.  

``` {r}
colnames(cleaned_loan)

```

``` {r}
usable_variables <- c("loan_amnt", "funded_amnt", "term", "int_rate", "installment", "emp_length", "home_ownership",
                        "annual_inc", "verification_status", "pymnt_plan", "purpose", "dti", "delinq_2yrs", 
                        "inq_last_6mths", "open_acc", "pub_rec", "revol_bal", "revol_util", "total_acc", 
                        "initial_list_status", "tot_coll_amt", "acc_now_delinq", "delinq_amnt", "tot_cur_bal", 
                        "avg_cur_bal", "collections_12_mths_ex_med", "chargeoff_within_12_mths", "bc_open_to_buy", 
                        "bc_util", "acc_open_past_24mths", "mo_sin_old_il_acct", "mo_sin_old_rev_tl_op", 
                        "mo_sin_rcnt_rev_tl_op", "mo_sin_rcnt_tl", "mths_since_recent_bc", "mths_since_recent_inq", 
                        "num_accts_ever_120_pd", "num_actv_bc_tl", "num_actv_rev_tl", "num_bc_sats", "num_bc_tl", 
                        "num_il_tl", "num_op_rev_tl", "num_rev_accts", "num_rev_tl_bal_gt_0", "num_sats", 
                        "num_tl_120dpd_2m", "num_tl_30dpd", "num_tl_90g_dpd_24m", "num_tl_op_past_12m", "pct_tl_nvr_dlq", 
                        "percent_bc_gt_75", "pub_rec_bankruptcies", "tax_liens", "tot_hi_cred_lim", "total_bal_ex_mort", 
                        "total_bc_limit", "total_il_high_credit_limit", "disbursement_method", "target")

cleaned_loan_2 <- cleaned_loan[,colnames(cleaned_loan) %in% usable_variables]
```

``` {r}
head(cleaned_loan_2)

```

### export the cleaned version of our data to a csv file and run model on it.

```{r}

write.csv(cleaned_loan_2, "cleaned_loan.csv")
```


